Tor's Usability for Censorship Circumvention

### 2015-12-11 paper/circumvention-experiment.tex
wordsmithing to get colums looking nice and killing orphans

There are still many features of Tor that are left unevaluated through user
research --- such as advanced web tasks (e.g., accessing hidden services), the
configuration menu to connect to the Tor network, configuring automatic
updates, and identity/cookie management. Rather than selecting the features to
study in isolation, we decided to focus on an important use case of Tor browser: censorship circumvention. 
Internet censorship is pervasive across the world today \cite{faris2008measuring}. 
Although Tor was not designed to be a censorship circumvention tool, users started
using it as one, since Tor's role as a proxy to anonymize
people also worked to circumvent censorship. Now, circumventing censorship is a
common use case for Tor, and current censorship circumvention methods are
incorporated into Tor, making it a viable censorship circumvention tool even in the face
of adversaries. In fact, some countries unsuccessfully attempt to block Tor for this reason \cite{winter2012great}. 
To our knowledge, this is the first user study investigating the usability of Tor as a 
censorship circumvention tool.

@@

\noindent {\bfseries Overview}
The Tor Browser provides a configuration interface to guide users through setting up 
their connection to the
Tor network. While people in most countries can connect to Tor using the
dialog's default settings, users in certain censorship environments must
make use of additional settings, such as bridges, proxies, or both. 
Our study explores how successfully this interface guides users to
correctly configuring their browser.

@@

We collected feedback from researchers, professors, and Tor users and developers
through the Tor project's mailing lists. We then iterated on the design by
incorporating the comments and suggestions we received into the next version of
the interface. We plan to take some qualitative measurements of the new design
by having some participants think aloud while they use our realistic mockup. This is 
to get any feedback before we launch our quantitative interface evaluation.

@@

We evaluated at least one Tor user, one participant who has heard of Tor before the study, and three people who had never heard of Tor before in each simulated censorship environment. The nature of small-scale qualitative studies results in rich information about each participant, but making generalizations from these results such as ``Tor users were able to configure quicker'' would be unsound. Our observations from our particular three participants were that they were more capable of debugging after failure modes and were more methodological about doing so, but they faced similar difficulties as the other participants.

We noticed four common challenges that our participants have had with the interface:

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Challenge 1:} People don't know how to choose between ``connect directly'' versus ``configure'' on the first screen. As seen in Figure \ref{fig:window1}, there is some guidance for the users, but we found this largely unhelpful since people do not understand the text on the screen. 
\item {\bfseries Challenge 2:} People feel compelled to set up a bridge, even if they don't need one, because they do not know how censorship works. Participants knew that they were censored to some degree since we had simulated censorship of websites. However, bridges are only necessary when the government has taken active steps to censor Tor itself.
\item {\bfseries Challenge 3:} When their first attempt to connect fails, people tend to first try changing the proxy settings, because the interface directs people to the last window participants saw before the connection failure. This was problematic since the problem is not necessarily where the interface directs users. For instance, it was common for the interface to redirect users to the proxy window in Figure \ref{fig:window5}, even when they needed to pick another bridge. 
\item {\bfseries Challenge 4:} When the first attempt to connect fails, people did not know what to do or what to try next. The feedback given to users was not understood (see Figure \ref{fig:error}) and also did not guide users into specific actions (such as trying a different bridge, trying the same configuration again, try connecting directly, etc.).
\end{itemize}

@@

These common four problems were illustrated by our participants concretely:

@@

A progress indicator at the top of the window indicates various steps necessary
for configuring a connection. In addition to communicating to the user what the various 
steps of configuration are from the beginning, this indicator also serves to clarify the independence of
bridges and proxies, which need to be configured separately.

In the event of a failure, the user is
returned to the first step of the process -- the bridge configuration -- to
avoid misleading users into thinking the proxy settings are the culprit. Previously,
the user had to manually click to go backward in the dialog to configure bridges.
In the rare event that the proxy is the issue for the connection setting, the user
will be directed to the proxy page naturally through the flow of the UI. \\

@@

The progress bar in the connection window drew many complaints, because it got
stuck and didn't offer any guidance for how long users were supposed to wait. 
The connection process can take a non-trivial amount of time, and users may
think that the connection isn't working, even when it is making slow progress. We
address this by providing a countdown timer, indicating that a long wait time does
not indicate failure. We further encourage waiting by adding visual indicators
for major milestones in the connection, providing better feedback about what has
been accomplished so far. Additionally, the current status of the connection
attempt (drawn from the log) is displayed, to aid in understanding and
debugging.

@@

In the event that the connection attempt fails, messages from the log are
displayed, along with tips and suggestions about what to do next (e.g., try
another bridge). To make the potential options clearer, the new failure screen
includes a separate button for common next steps which are able to resolve
a failed connection, which are to retry without changing settings and to reconfigure 
the connection. \\

@@

\noindent {\bfseries Additional changes}
In addition to the changes above, we have simplified flows where possible,
provided clearer definitions of bridges and proxies, and reduced visual clutter.
We believe that all of these changes are cohesive with the other design changes
made to the interface and will be beneficial for the user.

@@

The work summarized in the paper is only a part of the process of improving the Tor configuration
dialog for more usable censorship circumvention. We plan to continue this work by:

>>>>

Many features of Tor that are left unevaluated through user
research --- such as advanced web tasks (e.g., accessing hidden services), the
configuration menu to connect to the Tor network, configuring automatic
updates, and identity/cookie management. Rather than selecting the features to
study in isolation, we decided to focus on an important use case of Tor browser: censorship circumvention. 
Internet censorship is pervasive across the world today \cite{faris2008measuring}. 
Although Tor was not designed to be a censorship circumvention tool, users started
using it as one, since Tor's role as a proxy to anonymize
people also worked to circumvent censorship. Now, circumventing censorship is a
common use case for Tor, and current censorship circumvention methods are
incorporated into Tor, making it a viable censorship circumvention tool even in the face
of adversaries. In fact, some countries unsuccessfully attempt to block Tor for this reason \cite{winter2012great}. 
To our knowledge, this is the first user study investigating the usability of Tor as a 
censorship circumvention tool.

@@

\noindent {\bfseries Overview}
The Tor Browser provides a configuration interface to guide users through setting up 
their connection to the
Tor network. While most can connect to Tor using the
dialog's default settings, users in certain censorship environments must
make use of additional settings, such as bridges, proxies, or both. 
Our study explores how the average user circumvents censorship
through this interface.

@@

We collected feedback from researchers, professors, and Tor users and developers
through the Tor project's mailing lists. We iteratively designed by
incorporating the comments and suggestions we received into the next version of
the interface. We plan to take some qualitative measurements of the new design
by having some participants think aloud while they use our realistic mockup. This is 
to get any feedback before we launch our quantitative interface evaluation.

@@

In each censorship environment, we evaluated one Tor user, one who has heard of Tor before the study, and at least three people who had never heard of Tor before. Small-scale qualitative studies results in rich information about each participant, but making generalizations from these results such as ``Tor users were able to configure quicker'' would be unsound. Our observations of our particular three Tor users were that they were more capable of debugging after failure modes and were more methodological about doing so, but they faced similar difficulties as the other participants. \\

We noticed four common challenges:

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Challenge 1:} People don't know how to choose between ``connect directly'' versus ``configure'' on the first screen. As seen in Figure \ref{fig:window1}, there is some guidance for the users, but we found this largely unhelpful since people do not understand the text on the screen. 
\item {\bfseries Challenge 2:} People feel compelled to set up a bridge, even if they don't need one, because they do not know how censorship works. Participants knew that they were censored to some degree since we had simulated censorship of websites. However, bridges are only necessary when the government has taken active steps to censor Tor itself.
\item {\bfseries Challenge 3:} When their first attempt to connect fails, the interface directs people to the last window participants saw before the connection failure. The problem is not necessarily where the interface directs users, misguiding them. It was common for the interface to redirect users to the proxy window in Figure \ref{fig:window5}, even when they needed to pick another bridge. 
\item {\bfseries Challenge 4:} When the first attempt to connect fails, people did not know what to do or what to try next. The feedback given to users was not understood (see Figure \ref{fig:error}) and also did not guide users into specific actions (such as trying a different bridge, trying the same configuration again, try connecting directly, etc.).
\end{itemize}

@@

These problems were illustrated by our participants:

@@

A progress indicator at the top of indicates various steps
for configuring a connection. In addition to communicating what the various 
steps of configuration are from the beginning, this indicator also serves to clarify the independence of
bridges and proxies, which need to be configured separately.

After a failure, the users are
returned to the first step of the process -- the bridge configuration -- since this
is usually the cause for configuration issues. Previously,
the user had to manually click to go backward in the dialog to configure bridges.
In the rare event that the proxy is the issue for the connection setting, the user
will be directed to the proxy page naturally through the flow of the UI. \\

@@

The progress bar in the connection window often was unresponsive 
and did not offer guidance for to wait. 
The connection process can take about two minutes long enough that users may
think that the connection has failed, even when it is making slow and invisible progress. 
We added a countdown timer to indicate that a long wait time does
not indicate failure. We also add visual indicators
for major milestones in the connection, providing better feedback about what has
been accomplished so far. Additionally, the current status of the connection
attempt (drawn from the log) is displayed, to aid in understanding and
debugging.

@@

In the event that the connection attempt fails, messages from the log are
displayed, along with suggestions about what to do next (e.g., try
another bridge). To make the potential options clearer, the new failure screen
includes a separate button for common next steps which are able to resolve
a failed connection, which are to retry without changing settings and to reconfigure 
the connection. \\

@@

\noindent {\bfseries Additional changes}
In addition to the changes above, we have simplified flows,
reduced text, and reduced visual clutter.
We believe these changes are cohesive with the our design 
and will be beneficial for the user.\\

@@

The work summarized is only a part of the process of improving the Tor configuration
dialog for more usable censorship circumvention. We plan to continue this work by:
